\section{Implementation}

\subsection{Picking a programming language}
We decided to program AiryScript in Scala fairly early on for the following
reasons.
\begin{itemize}
  \item We wanted to be able to really define our own language syntax.
    Generally, this means having our own parser and not merely calling 'eval' in
    some other language. Also, all of the members of our team prefer typed
    languages. Therefore, we also wanted to create a typed language. We could
    not see how to do this without making our own parser. We were already in
    possession of some basic parser code for Scala.
  \item Some of the members of our team were already familiar with Scala.
  \item The whole team has a thorough knowledge in Java gained in other courses
  and projects. This smooths the learning curve of learning (the basic
  principles of) an almost entirely new language. 
  \item Scala has very good support for pattern matching through case classes
    (i.e. tagged unions). We make extensive use of this in our code.
  \item Some of the members of our team had been looking for a good opportunity
    to get to know Scala for some time.
\end{itemize}


\subsection{Implementation overview}

We divided the implementation in 3 parts:
\begin{enumerate}
\item The parser: Responsible for parsing a string into the correct operation,
or returning a parse error if the given expression was incorrect.
\item The operation handler: This component contains the implementation logic
of all the operators. The handler is responsible for the consistency checking of
all the provided arguments of an operation and either throwing an error if one
or more arguments were incorrect/incomplete or performing the operation. The
database will be accessed for verification as well as performing the actual
\textsc{crud} operation.
\item The database: We used an SQL database to persist. The
implementation of this database is closely related to the domain model.
\end{enumerate} 
	
\par
All of the operations and types are represented by a single case class in Scala,
which captures all of the required information to use this. All three of the
aforementioned components are united in a single script which will execute them
in the following order. First, the parser will parse the user's input into an
operation, after which this is passed to the second phase. The handler will
subssequently handle these operations by pattern matching over the different
kinds of operators and their respective arguments. In case an error occured
somewhere along this execution path, the user will be informed of what went wrong.


\input{4-implementation/parser}
\input{4-implementation/handler} 
\input{4-implementation/database}
